Recovering from population extinction in the Animal Life Cycle Algorithm (ALCA)

### 2022-09-23 paper.tex
Updated Exeriments
Work in progress.

The alternatives to restart this new population are the creation of a new set
of candidate solutions based on either: The mutation with uniform modification
from the Elite or the historically best-found solution (Champion); Crossing the
Champion with the Elite; The projection of the Elite towards the Champion;
Based on random use of the previous alternatives.

@@

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% To change the size of the font use \scriptsize or \tiny
\begin{table}[]
    \scriptsize
    \centering        
    \caption{Animal Life Cycle Algorithm (ALCA) General Configuration.}\label{tab.general_configuration}
    \begin{tabular}{@{}ll@{}}
    \toprule
    \multicolumn{2}{l}{ALCA General Configuration} \\ \midrule
    Function name & ALL \\
    Population & 500 \\
    Target error & 1.00E-08 \\
    Crossover rate & 100 \\
    Mutation rate & 7 \\
    Max age & 40 \\
    Tournament rep. & 100 \\
    Sample size & 20 \\
    Base approval & 80 \\
    Goal approval & 200 \\ \bottomrule
    \end{tabular}
    \end{table}

\begin{table}[]
    \scriptsize
    \centering        
    \caption{Classic Benchmark Functions for Optimization Configuration.}\label{tab.benchmark_fun_config}
    \begin{tabular}{@{}llllllll@{}}
    \toprule
    \multicolumn{8}{l}{Benchmark Configuration Setup} \\ \midrule
    Function name & Ackley & Bohachevsky & Griewank & Rastrigin & Sphere & Rosenbrock & Rosenbrock \\
    Dimensions & 5, 10 & 5, 10 & 5, 10 & 5, 10 & 5, 10 & 5 & 10 \\
    Max evaluations & 200,000 & 200,000 & 200,000 & 200,000 & 200,000 & 500,000 & 1,000,000 \\
    Max stagnation & 50,000 & 50,000 & 50,000 & 50,000 & 50,000 & 125,000 & 250,000 \\
    Target fitness & -20.0 & -12.0 & -150.0 & -80.0 & -50.0 & -3500.0 & -3500.0 \\
    Bound matrix & {[}-32, 32{]} & {[}-2, 2{]} & {[}-500, 500{]} & {[}-5, 5{]} & {[}-5, 5{]} & {[}-2, 2{]} & {[}-2, 2{]} \\ \bottomrule
    \end{tabular}
    \end{table}

@@

>>>>

@@

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% To change the size of the font use \scriptsize or \tiny
\begin{table}[]
    \scriptsize
    \centering
    \caption{Animal Life Cycle Algorithm (ALCA) General Configuration.}\label{tab.general_configuration}
    \begin{tabular}{@{}ll@{}}
    \toprule
    \multicolumn{2}{l}{\textbf{ALCA General Configuration}} \\ \midrule
    Function name & ALL \\
    Population & 500 \\
    Target error & 1.00E-08 \\
    Crossover rate & 100 \\
    Mutation rate & 7 \\
    Max age & 40 \\
    Tournament rep. & 100 \\
    Sample size & 20 \\
    Base approval & 80 \\
    Goal approval & 200 \\ \bottomrule
    \end{tabular}
    \end{table}

\begin{table}[]
    \scriptsize
    \centering
    \caption{Classic Benchmark Functions for Optimization Configuration.}\label{tab.benchmark_fun_config}
    \begin{tabular}{@{}llllllll@{}}
    \toprule
    \multicolumn{8}{l}{\textbf{Benchmark Configuration Setup}} \\ \midrule
    Function name & Ackley & Bohachevsky & Griewank & Rastrigin & Sphere & Rosenbrock & Rosenbrock \\
    Dimensions & 5, 10 & 5, 10 & 5, 10 & 5, 10 & 5, 10 & 5 & 10 \\
    Max evaluations & 200,000 & 200,000 & 200,000 & 200,000 & 200,000 & 500,000 & 1,000,000 \\
    Max stagnation & 50,000 & 50,000 & 50,000 & 50,000 & 50,000 & 125,000 & 250,000 \\
    Target fitness & -20.0 & -12.0 & -150.0 & -80.0 & -50.0 & -3500.0 & -3500.0 \\
    Bound matrix & {[}-32, 32{]} & {[}-2, 2{]} & {[}-500, 500{]} & {[}-5, 5{]} & {[}-5, 5{]} & {[}-2, 2{]} & {[}-2, 2{]} \\ \bottomrule
    \end{tabular}
    \end{table}

@@

For each experiment, we ran 30 independent executions per algorithm and
dimensions specified. We recorded the following results: best found error,
standard deviation, total number of evaluations, and total elapsed time (in
seconds). The labels used on the results of our summarized experiments tables
are the following:

\begin{itemize}
    \item   Error:       is the best found error (or best solution).
    \item   St-Dev:      is the standard deviation calculated from the error. 
    \item   Evals.:      is the total number of evaluations the algorithm executed.
    \item   Time:        is the total elapsed or wall clock time, in seconds. 
\end{itemize}

The alternatives to restart this new population are the creation of a new set
of candidate solutions based on either: 1. Crossing the Champion with the Elite;
2. The mutation with uniform modification from the Elite; 3. The
projection of the Elite towards the Champion; 4. Based on random use of the
previous alternatives. We compared three strategies by grouping the previous
alternatives in the following:

\begin{itemize}
    \item   Xover Mutation:      Crossing the Champion with the Elite, and the mutation with uniform modification from the Elite.
    \item   Elite Projection:    The projection of the Elite towards the Champion. 
    \item   All Random:          Based on random use of the previous alternatives.
\end{itemize}

\subsubsection{Ackley Benchmark Function Experiment}

\begin{figure}
    \includegraphics[width=\textwidth]{img/fig_fun_ackley.pdf}
    \caption{Box and whisker chart for the Ackley Benchmark Function, 5 vs 10 Dimensions comparison.} \label{fig.fun_ackley}
    \end{figure}

\begin{table}[]
    \scriptsize
    \centering
    \caption{Ackley Benchmark Function summarized experiments table for 5 Dimensions.}\label{tab.benchmark_fun_ackley5}
    \begin{tabular}{@{}lllll@{}}
    \toprule
    \multicolumn{5}{l}{\textbf{Ackley Function - 5 Dimensions}} \\ \midrule
     & \textbf{Error} & \textbf{St-Dev} & \textbf{Evals.} & \textbf{Time} \\
    \textbf{Xover Mutation} & 8.69E-08 & 1.36E-08 & 46,331 & 49.9 \\
    \textbf{Elite Projection} & 6.33E-08 & 2.70E-08 & 13,811 & 18.1 \\
    \textbf{All Random} & 6.78E-08 & 2.55E-08 & 30,236 & 32.8 \\ \bottomrule
    \end{tabular}
    \end{table}

\begin{table}[]
    \scriptsize
    \centering
    \caption{Ackley Benchmark Function summarized experiments table for 10 Dimensions.}\label{tab.benchmark_fun_ackley10}
    \begin{tabular}{@{}lllll@{}}
    \toprule
    \multicolumn{5}{l}{\textbf{Ackley Function - 10 Dimensions}} \\ \midrule
     & \textbf{Error} & \textbf{St-Dev} & \textbf{Evals.} & \textbf{Time} \\
    \textbf{Xover Mutation} & 9.35E-08 & 5.93E-09 & 104,842 & 111.7 \\
    \textbf{Elite Projection} & 5.11E-08 & 3.13E-08 & 16,019 & 20.1 \\
    \textbf{All Random} & 6.89E-08 & 2.75E-08 & 43,631 & 46.3 \\ \bottomrule
    \end{tabular}
    \end{table}
